---
title: "Untitled"
author: "John Cornwell"
date: "January 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
The purpose of this script is to generate some preliminary stats and plots to look at variation in usage among end uses in the NEEA RBSA data to guide the EULR study sampling.

## call packages.
```{r , include=FALSE}
#Package Setup
x <- c("ggmap", "rgdal", "rgeos", "maptools","plyr", "dplyr", "tidyr", "reshape2", "xlsx", "XLConnect", "readxl", "lubridate", "data.table", "stringr")
#install.packages(x) # warning: uncommenting this may take a number of minutes
lapply(x, library, character.only = TRUE) # load the required packages
```

## laod full data
```{r}
#Import Files
meter_data <- fread("RBSA_METER_DATA.csv", header = T, stringsAsFactors = F)
# meter_er <- fread("RBSA_METER_DATA.csv", header = T, stringsAsFactors = F, select = c('siteid', 'time', 'ER', "YEAR", "MONTH", "DAY", "HOUR", "MINUTE", "WEEK", "WEEKEND", "WEEKDAY")) 
## find homes with ER data and filter
# meter_er_summary_A <- meter_er %>% 
#   group_by(siteid) %>%
#   summarise(sum_er = sum(ER))
# 
# meter_er_summary <- subset(meter_er_summary_A, !is.na(meter_er_summary$sum_er))
# meter_er_filter <- subset(meter_er, meter_er$siteid %in% meter_er_summary$siteid)

meter_data_y2 <- fread("RBSAM Year 2.txt", header = T, stringsAsFactors = F)
sitedata1 <- read.xlsx("SFMaster_housegeometry.xlsx", sheetIndex = 1) 
sitedata2 <- read.xlsx("SFMaster_location.xlsx", sheetIndex = 1) 
sitedata3 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFMaster_populations.xlsx", sheetIndex = 1)
sitedata4 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFwheater.xlsx", sheetIndex = 1)
sitedata5 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SF_ri_demog.xlsx", sheetIndex = 1)
sitedata6 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SF_ri_heu.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFshowerhead.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/HVACcooling.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/HVACheating.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFblowerdoor.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFatticceiling.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFducts.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFflbasement.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFflcrawl.xlsx", sheetIndex = 1)
sitedata7 <- read.xlsx("2011 RBSA Single Family Database (Excel)/SFframedwall.xlsx", sheetIndex = 1)



```

## remove columns we don't need
```{r}
cols_keep <- names(meter_data)
cols_remove <- c(4:11, 13:15, 17:35, 39:40, 43, 46:56, 58:73, 77:95, 97:139, 141:156, 158:162, 164:176, 180:189, 192:193, 195:199, 202:222)
cols_keep <- cols_keep[-cols_remove]
meter_data2 <- as.data.frame(meter_data)
meter_data2 <- meter_data2[ ,cols_keep]

### gather columns into rows with measure names
meter_data3 <- gather(meter_data2, key = "measure", value = "kw", 3:28)
meter_data3 <- subset(meter_data3, !is.na(meter_data3$kw))
meter_data3$date <- dmy_hms(meter_data3$time)
meter_data3$date2 <- as_date(meter_data3$date)
meter_data3$hour <- hour(meter_data3$date)

meter_data3 <- data.table(meter_data3)
## extract temp data and get mean hourly from 15 minute data for each home / measure / hour .
meter_data_temp <- meter_data3[meter_data3$measure == "ODT" | meter_data3$measure == "IDT" | meter_data3$measure == "WST" , ]
meter_data_temp_h <- meter_data_temp[,list(temp=mean(kw)),by=list(siteid,measure,date2,hour), verbose = TRUE] 
meter_data_temp <- spread(meter_data_temp_h, measure, temp)
rm(meter_data_temp_h)

## sum hourly from 15 minute data for each home / measure / hour .
meter_data3 <- meter_data3[meter_data3$measure != "ODT" & meter_data3$measure != "IDT" & meter_data3$measure != "WST" , ]
meter_data3_h <- meter_data3[,list(kw_h=sum(kw)),by=list(siteid,measure,date2,hour), verbose = TRUE] 
rm(meter_data3)

## spread data so there is one row per hour with variables for measures
meter_data_wide <- spread(meter_data3_h, measure, kw_h)
rm(meter_data3_h)
## combine measure and temp data
meter_data_final <- left_join(meter_data_wide, meter_data_temp, by = c("siteid", "date2", "hour"))
rm(meter_data_temp, meter_data_wide, meter_data2)
write.csv(meter_data_final, "meter_data_final.csv")
rm(meter_data, meter_data_final)
## add in site characteristics data


## create season flags
meter_data3_h$season <-  ifelse(month(meter_data3_h$date2) > 5 & month(meter_data3_h$date2) < 9, "SUMMER",NA)
meter_data3_h$season <-  ifelse(month(meter_data3_h$date2) > 8 & month(meter_data3_h$date2) < 12, "FALL",meter_data3_h$season)
meter_data3_h$season <-  ifelse(month(meter_data3_h$date2) == 12 | month(meter_data3_h$date2) < 3, "WINTER",meter_data3_h$season)
meter_data3_h$season <-  ifelse(month(meter_data3_h$date2) > 2 & month(meter_data3_h$date2) < 6, "SPRING",meter_data3_h$season)

## create datasets for aggregate daily loadshapes per house for each season and all year.

meter_data3_h_an <- meter_data3_h[,list(kw_h=mean(kw_h)),by=list(siteid, measure,hour), verbose = TRUE] 
meter_data3_h_season <- meter_data3_h[,list(kw_h=mean(kw_h)),by=list(siteid, measure, season, hour), verbose = TRUE] 
meter_data3_h_an$season <- NA

## create master summary dataset that includes hourly load shape data for each period/site/measure. periods are annual, summer, spring, fall, winter.

meterdata_summary_hrs <- rbind(meter_data3_h_an, meter_data3_h_season)
meterdata_summary_hrs$season[is.na(meterdata_summary_hrs$season)] <- "ANNUAL"

## create datasets with mean daily total use by period / site / measure.
meterdata_summary_day <- meterdata_summary_hrs[,list(kwh=sum(kw_h)), by=list(siteid,season, measure), verbose = TRUE] 

```

#Now have two datasets 
1) meterdata_summary_hrs - Average daily load shape (24 hour observations, per season, site, measure. 
2) meterdata_summary_day - Average daily total by season, site, measure. 

#For now we are only interested in 6 measure types: DHP, Ducted HEat Pump, Water heaters, electric baseboard heat, electric furnace, and CAC. Also included PTAC and PTHP. For this analysis we will select the columns for these rows only along with weather and total consumption variable. 

```{r}
## select measures of interest
keep <- c("AC", "DHP", "DHW_1", "DHW_2", "DHW_HP", "ER", "ER_2", "ER_3", "ER_4", "ER_5", "GSHP_in", "GSHP_out", "HP_in", "HP_in_2", "HP_out", "HP_out_2", "HP_VLT", "IDT", "ODT", "Service", "WST", "Furn", "PTAC", "PTHP")

meterdata_summary_hrs_analysis <- meterdata_summary_hrs[meterdata_summary_hrs$measure %in% keep, ]
meterdata_summary_day_analysis <- meterdata_summary_day[meterdata_summary_day$measure %in% keep, ]    

## sum values for DHW, ER, GSHP, HP, for hourly database
meterdata_summary_hrs_analysis$measure2 <- NA
meterdata_summary_hrs_analysis$measure2 <- ifelse(meterdata_summary_hrs_analysis$measure == "DHW_1" | meterdata_summary_hrs_analysis$measure == "DHW_2", "DHW", meterdata_summary_hrs_analysis$measure2)
meterdata_summary_hrs_analysis$measure2 <- ifelse(str_detect(meterdata_summary_hrs_analysis$measure, "ER") == TRUE, "ER", meterdata_summary_hrs_analysis$measure2)
meterdata_summary_hrs_analysis$measure2 <- ifelse(str_detect(meterdata_summary_hrs_analysis$measure, "GSHP") == TRUE, "GSHP", meterdata_summary_hrs_analysis$measure2)
meterdata_summary_hrs_analysis$measure2 <- ifelse(meterdata_summary_hrs_analysis$measure == "HP_in" | meterdata_summary_hrs_analysis$measure == "HP_in_2"| meterdata_summary_hrs_analysis$measure == "HP_out"| meterdata_summary_hrs_analysis$measure == "HP_out_2", "HP", meterdata_summary_hrs_analysis$measure2)
meterdata_summary_hrs_analysis$measure2 <- ifelse(is.na(meterdata_summary_hrs_analysis$measure2),meterdata_summary_hrs_analysis$measure , meterdata_summary_hrs_analysis$measure2)
meterdata_summary_hrs_analysis <- meterdata_summary_hrs_analysis[,list(kw_h2=sum(kw_h)),by=list(siteid,season,measure2,hour), verbose = TRUE] 

## sum values for DHW, ER, GSHP, HP, for daily database
meterdata_summary_day_analysis$measure2 <- NA
meterdata_summary_day_analysis$measure2 <- ifelse(meterdata_summary_day_analysis$measure == "DHW_1" | meterdata_summary_day_analysis$measure == "DHW_2", "DHW", meterdata_summary_day_analysis$measure2)
meterdata_summary_day_analysis$measure2 <- ifelse(str_detect(meterdata_summary_day_analysis$measure, "ER") == TRUE, "ER", meterdata_summary_day_analysis$measure2)
meterdata_summary_day_analysis$measure2 <- ifelse(str_detect(meterdata_summary_day_analysis$measure, "GSHP") == TRUE, "GSHP", meterdata_summary_day_analysis$measure2)
meterdata_summary_day_analysis$measure2 <- ifelse(meterdata_summary_day_analysis$measure == "HP_in" | meterdata_summary_day_analysis$measure == "HP_in_2"| meterdata_summary_day_analysis$measure == "HP_out"| meterdata_summary_day_analysis$measure == "HP_out_2", "HP", meterdata_summary_day_analysis$measure2)
meterdata_summary_day_analysis$measure2 <- ifelse(is.na(meterdata_summary_day_analysis$measure2),meterdata_summary_day_analysis$measure , meterdata_summary_day_analysis$measure2)
meterdata_summary_day_analysis <- meterdata_summary_day_analysis[,list(kwh2=sum(kwh)),by=list(siteid,season,measure2), verbose = TRUE] 

```

# Scale daily dataset values to z scores
```{r}
meterdata_summary_day_analysis_2 <- meterdata_summary_day_analysis %>% 
                                    group_by(season, measure2)  %>%
                                    mutate(kwh_z = scale(kwh2, center = T, scale = T))
meterdata_summary_day_analysis_2 <- ungroup(meterdata_summary_day_analysis_2)

## test plot to confirm scaleing worked.
test <- filter(meterdata_summary_day_analysis_2, measure2 == "ER")

ggplot(test[test$season == "WINTER", ], aes(x=kwh2, y=kwh_z)) +
  geom_point()

mean(test[test$season == "WINTER", ]$kwh2)
```

## Plots

# Plot 1: zscore distribution charts
```{r}
keep <- c("AC", "DHP", "DHW", "ER", "HP", "Furn")
zplotdata <- meterdata_summary_day_analysis_2[meterdata_summary_day_analysis_2$measure2 %in% keep, ]

zplotdata$measure2 <- str_replace(zplotdata$measure2, "AC", "Central AC")
zplotdata$measure2 <- str_replace(zplotdata$measure2, "DHP", "Ductless Heat Pump")
zplotdata$measure2 <- str_replace(zplotdata$measure2, "DHW", "Hot Water Heater")
zplotdata$measure2 <- str_replace(zplotdata$measure2, "ER", "Electric Baseboard Heat")
zplotdata$measure2 <- str_replace(zplotdata$measure2, "HP", "Ducted Heat Pump")
zplotdata$measure2 <- str_replace(zplotdata$measure2, "Furn", "Electric Furnace")
colnames(zplotdata)[3] <- "Measure"
colnames(zplotdata)[5] <- "Z_Score"

#boxplot
box <- ggplot(zplotdata, aes(x = Measure, y = Z_Score, fill = Measure)) +
      geom_boxplot() +
      stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y..), width = .75, linetype = "dashed") + #dashed line at mean
      coord_flip() +
      theme_classic() +
      theme(plot.title = element_text(hjust = .5), axis.ticks = element_blank(), legend.position = "none") +   # Centre plot title
      scale_fill_brewer(palette = "YlGn") + # Color palette 
    
ggsave("box.pdf", width = 8, height = 5, units = "in", dpi = 1000)

box2 <- ggplot(zplotdata[zplotdata$season != "ANNUAL", ], aes(x = Measure, y = Z_Score, fill = Measure)) +
      geom_boxplot() +
      coord_flip() +
      theme_classic() +
     theme(plot.title = element_text(hjust = .5), 
                                    axis.ticks = element_blank(), legend.position = "none") +   # Centre plot title
                              scale_fill_brewer(palette = "YlGn")  # Color palette
box2 + facet_wrap( ~ season, ncol=2)
ggsave("box_season.pdf", width = 8, height = 5, units = "in", dpi = 1000)

#violin plot
ggplot(zplotdata, aes(x = Measure, y = Z_Score, fill = Measure)) +
      geom_violin() +
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y..), width = .75, linetype = "dashed") + #dashed line at mean
      coord_flip() +
      theme_classic() +
     theme(plot.title = element_text(hjust = .5), 
                                    axis.ticks = element_blank(), legend.position = "none") +   # Centre plot title
                              scale_fill_brewer(palette = "YlGn")  # Color palette
ggsave("violin.pdf", width = 8, height = 5, units = "in", dpi = 1000)

v <- ggplot(zplotdata[zplotdata$season != "ANNUAL", ], aes(x = Measure, y = Z_Score, fill = Measure)) +
      geom_violin() +
  stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y..), width = .75, linetype = "dashed") + #dashed line at mean
      coord_flip() +
      theme_classic() +
     theme(plot.title = element_text(hjust = .5), 
                                    axis.ticks = element_blank(), legend.position = "none") +   # Centre plot title
                              scale_fill_brewer(palette = "YlGn")  # Color palette
v + facet_wrap( ~ season, ncol=2)
ggsave("violin.pdf", width = 8, height = 5, units = "in", dpi = 1000)

```

```{r pressure, echo=FALSE}

meter_er_winter <- subset(meter_er_filter, meter_er_filter$MONTH > 9 | meter_er_filter$MONTH < 4)
meter_er_w_day_load <- meter_er_winter %>%
                        group_by(siteid, HOUR) %>%
                        summarise(mean_h = mean(ER))

meter_er_w_day_load_s <- meter_er_w_day_load %>% 
                          group_by(HOUR) %>%
                          summarise(mean_h_s = mean(mean_h, na.rm = T),
                                    sd_h_s = sd(mean_h, na.rm = T),
                                    min_h_s = min(mean_h),
                                    max_h_s = max(mean_h))
meter_er_w_day_load_s$upper_sd <- meter_er_w_day_load_s$mean_h_s + meter_er_w_day_load_s$sd_h_s
meter_er_w_day_load_s$lower_sd <- meter_er_w_day_load_s$mean_h_s - meter_er_w_day_load_s$sd_h_s


ER_plotdata <- meterdata_summary_hrs_analysis[meterdata_summary_hrs_analysis$measure2 == "ER" & meterdata_summary_hrs_analysis$season == "ANNUAL", ]

ggplot(ER_plotdata) +
  geom_line(aes(x=hour, y=kw_h2), size = 2) +
  geom_ribbon(aes(x=hour, ymax = kw_h2 + sd_h_s, ymin = mean_h_s - sd_h_s, alpha = 0.2)) +
  geom_line(aes(x=hour, y=kw_h2)) +
  geom_line(aes(x=hour, y=kw_h2)) +
  coord_cartesian(ylim=c(0, .8))

```
# Start with baseborad 
```{r}
#Import Files
meter_ac <- fread("RBSA_METER_DATA.csv", header = T, stringsAsFactors = F, select = c('siteid', 'time', 'AC', "YEAR", "MONTH", "DAY", "HOUR", "MINUTE", "WEEK", "WEEKEND", "WEEKDAY")) 
## find homes with ER data and filter
meter_ac_summary_A <- meter_ac %>% 
  group_by(siteid) %>%
  summarise(sum_ac = sum(AC))

meter_ac_summary <- subset(meter_ac_summary_A, !is.na(meter_ac_summary_A$sum_ac))
meter_ac_filter <- subset(meter_ac, meter_ac$siteid %in% meter_ac_summary$siteid)
```

```{r}
# Analysis
# select winter months (Oct-MAR)

meter_ac_summer <- subset(meter_ac_filter, meter_ac_filter$MONTH > 3 & meter_ac_filter$MONTH < 10)
meter_ac_w_day_load <- meter_ac_summer %>%
                        group_by(siteid, HOUR) %>%
                        summarise(mean_h = mean(AC))

meter_ac_w_day_load_s <- meter_ac_w_day_load %>% 
                          group_by(HOUR) %>%
                          summarise(mean_h_s = mean(mean_h, na.rm = T),
                                    sd_h_s = sd(mean_h, na.rm = T),
                                    min_h_s = min(mean_h),
                                    max_h_s = max(mean_h))
meter_ac_w_day_load_s$upper_sd <- meter_ac_w_day_load_s$mean_h_s + meter_ac_w_day_load_s$sd_h_s
meter_ac_w_day_load_s$lower_sd <- meter_ac_w_day_load_s$mean_h_s - meter_ac_w_day_load_s$sd_h_s

meter_day_sum <- meter_ac_summer %>%
                  group_by(siteid, YEAR, MONTH, DAY) %>%
                  summarise(daysum_ac = sum(AC))

meter_avg_day <- meter_day_sum %>%
                  group_by(siteid) %>%
                  summarise(daily_ac = mean(daysum_ac))

meter_avg_day$scaled <- scale(meter_avg_day$daily_ac, center = T, scale = F)
meter_avg_day$ac_z <- scale(meter_avg_day$scaled, center = T, scale = T)

z_table <- merge(z_table, meter_avg_day[ , c("siteid", "ac_z")], by = "siteid", all.x = T )


```

``` {r}
```
